### 2015-02-24 paper.tex
add paper skeleton with abstract

\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{url}
\usepackage{subfigure}
\usepackage{booktabs,threeparttable,multirow}

% new math operators
\DeclareMathOperator{\abs}{abs}

% todo command
\usepackage{marginnote}
\newcounter{todocnt}
\newcommand{\Sim}{\textsc{Simon}} 
\setcounter{todocnt}{0}
\newcommand{\todo}[1]{\stepcounter{todocnt}{\tt {[#1]}} \marginpar{{$\blacksquare$ \thetodocnt}}}  
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\hyphenation{op-tical net-works semi-conduc-tor}
\IEEEoverridecommandlockouts
\begin{document}

\title{Countermeasures against co-location attempts in the cloud}

\author{\IEEEauthorblockN{Dolan Murvihill, Nathan Wells
}
\IEEEauthorblockA{Worcester Polytechnic Institute, 
Worcester, MA 01609, USA\\
Email: \texttt{\{dm, nhwells\}@wpi.edu}
}}
\maketitle

\begin{abstract}
Recent research has demonstrated that cloud services such as Amazon AWS do not
provide perfect isolation --- that there are side channels between virtual
machines that reside on the same host. In particular, it is possible, under
some circumstances, to use cache timing information to perform a full AES key
recovery against a VM sharing the same cache.

Other work has shown that cloud services such as Amazon AWS provision virtual
machines using straightforward, often predictable algorithms, and others have
demonstrated that it is possible to achieve the required co-location to perform
a co-location attack. We would like to interrupt this step in the kill chain,
by developing a provisioning strategy that makes co-location much more
difficult to force or predict, without sacrificing too much performance.
\end{abstract}

\section{Motivation}
\todo{Know what you want to do and why that is interesting (maybe with bullet points). But do not write this section until you know what you actually have done so that the motivation fits your work.}

To generate this pdf, you need a latex implementation. I recommend Texniccenter in Windows, but there are others. For your references (the .bib file) jabref is the only helpful editor I have used.
Refer to sections like this: Section~\ref{sec:background} and to references like this: Ristenpart et al.~\cite{Ristenpart_hey}. \\{\tiny There is also a word template available.}

We will stick to the following timeline:

\begin{itemize}
	\item 2/22: Project goals and outline defined
	\item 3/1 : Related work identified and described in report.
	\item 3/15: At least 1/3 of your anticipated work should be completed and documented
	\item 3/22: At least 2/3 of your anticipated work should be completed and documented
	\item 3/29: All of your anticipated work should be completed and documented
	\item 4/5:  Your results and outcomes are now also documented and included in the paper
	\item 4/12: Final submission of of complete paper for review
	\item 4/21: Reviews complete
	\item 4/26: Submission of final version addressing reviewers comments.
	\item 5/4 : Presentation of your project in class.
\end{itemize}

\section{Background}\label{sec:background}
\todo{You should find and describe related work early on. Know what other people have done.}

\section{Work Description}
Here you describe the work you have performed, problems you have solved and methods you have used. There is a fine balance between brevity and conciseness and ensuring that other people, if investing the time, would be able to reproduce your results given this description.

\section{Results}
\todo{here you will present and discuss your outcomes: implementation results or measurements or other project outcomes}

\section{Conclusion}
\todo{TBD last}

%\bibliographystyle{IEEEtran}

\end{document}

### 2015-02-25 paper.tex
fix to one sentence per line
Added Motivation

@@
 \begin{abstract}
-Recent research has demonstrated that cloud services such as Amazon AWS do not
-provide perfect isolation --- that there are side channels between virtual
-machines that reside on the same host. In particular, it is possible, under
-some circumstances, to use cache timing information to perform a full AES key
+Recent research has demonstrated that cloud services such as Amazon AWS do not provide perfect isolation --- that there
+  are side channels between virtual machines that reside on the same host.
+In particular, it is possible, under some circumstances, to use cache timing information to perform a full AES key
   recovery against a VM sharing the same cache.
 
-Other work has shown that cloud services such as Amazon AWS provision virtual
-machines using straightforward, often predictable algorithms, and others have
-demonstrated that it is possible to achieve the required co-location to perform
-a co-location attack. We would like to interrupt this step in the kill chain,
-by developing a provisioning strategy that makes co-location much more
-difficult to force or predict, without sacrificing too much performance.
+Other work has shown that cloud services such as Amazon AWS provision virtual machines using straightforward, often
+  predictable algorithms, and others have demonstrated that it is possible to achieve the required co-location to
+  perform a co-location attack.
+We would like to interrupt this step in the kill chain, by developing a provisioning strategy that makes co-location
+  much more difficult to force or predict, without sacrificing too much performance.
 \end{abstract}
@@
 
-To generate this pdf, you need a latex implementation. I recommend Texniccenter in Windows, but there are others. For your references (the .bib file) jabref is the only helpful editor I have used.
-Refer to sections like this: Section~\ref{sec:background} and to references like this: Ristenpart et al.~\cite{Ristenpart_hey}. \\{\tiny There is also a word template available.}
+\begin{itemize}
+	\item Amazon's EC2 service is very popular and used for by variety of customers, some with sensitive data
+	\item This makes VMs on this service a worthwhile target
+	\item Ristenpart et. al have demonstrated that it is possible to create a VM co-resident with a target instance and
+	that existing side-channel attacks can be employed against co-resident instances on the cloud
+	\item We would like to focus on a method to obscure the algorithm by which instances are placed, which would make it
+	significantly harder to achieve co-location
+	\item Alternatively, we could focus on a method which prevents the simple network-based checks used to determine if two instances are co-located.
 
-We will stick to the following timeline:
-
-\begin{itemize}
-	\item 2/22: Project goals and outline defined
-	\item 3/1 : Related work identified and described in report.
-	\item 3/15: At least 1/3 of your anticipated work should be completed and documented
-	\item 3/22: At least 2/3 of your anticipated work should be completed and documented
-	\item 3/29: All of your anticipated work should be completed and documented
-	\item 4/5:  Your results and outcomes are now also documented and included in the paper
-	\item 4/12: Final submission of of complete paper for review
-	\item 4/21: Reviews complete
-	\item 4/26: Submission of final version addressing reviewers comments.
-	\item 5/4 : Presentation of your project in class.
 \end{itemize}

### 2015-02-25 paper.tex
style fixes

@@
 \begin{abstract}
-Recent research has demonstrated that cloud services such as Amazon AWS do not provide perfect isolation --- that there
+Recent research has demonstrated that cloud services such as Amazon AWS do not provide perfect isolation~--- that there
   are side channels between virtual machines that reside on the same host.
@@
 \begin{itemize}
-	\item Amazon's EC2 service is very popular and used for by variety of customers, some with sensitive data
-	\item This makes VMs on this service a worthwhile target
+  \item Amazon's EC2 service is very popular and used for by variety of customers, some with sensitive data.
+  \item This makes VMs on this service a worthwhile target.
   \item Ristenpart et. al have demonstrated that it is possible to create a VM co-resident with a target instance and
-	that existing side-channel attacks can be employed against co-resident instances on the cloud
+    that existing side-channel attacks can be employed against co-resident instances on the cloud.
   \item We would like to focus on a method to obscure the algorithm by which instances are placed, which would make it
-	significantly harder to achieve co-location
-	\item Alternatively, we could focus on a method which prevents the simple network-based checks used to determine if two instances are co-located.
-
+    significantly harder to achieve co-location.
+  \item Alternatively, we could focus on a method which prevents the simple network-based checks used to determine if
+    two instances are co-located.
 \end{itemize}

### 2015-02-25 paper.tex
small style fix

@@
 \begin{itemize}
-  \item Amazon's EC2 service is very popular and used for by variety of customers, some with sensitive data.
-  \item This makes VMs on this service a worthwhile target.
+  \item Amazon's EC2 service is very popular and used for by variety of customers, some of which have sensitive data.
+  \item This makes VMs on EC2 a worthwhile target.
   \item Ristenpart et. al have demonstrated that it is possible to create a VM co-resident with a target instance and

### 2015-03-18 paper.tex
move shared preamble code into preamble.tex
add work plan

@@
 
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{op-tical net-works semi-con-duc-tor}
 \IEEEoverridecommandlockouts
@@
   perform a co-location attack.
-We would like to interrupt this step in the kill chain, by developing a provisioning strategy that makes co-location
-  much more difficult to force or predict, without sacrificing too much performance.
+We would like to discern whether Amazon AWS has implemented any countermeasures against co-location attacks since the
+first successful cloud co-location six years ago; we will repeat their procedure in an attempt to reproduce their
+results. If time permits, we will try to improve on their attack.
 \end{abstract}
@@
   \item This makes VMs on EC2 a worthwhile target.
-  \item Ristenpart et. al have demonstrated that it is possible to create a VM co-resident with a target instance and
+  \item Ristenpart et al. have demonstrated that it is possible to create a VM co-resident with a target instance and
     that existing side-channel attacks can be employed against co-resident instances on the cloud.
-  \item We would like to focus on a method to obscure the algorithm by which instances are placed, which would make it
-    significantly harder to achieve co-location.
-  \item Alternatively, we could focus on a method which prevents the simple network-based checks used to determine if
-    two instances are co-located.
+  \item We expect that Amazon has since adjusted their VM location algorithms to make the Ristenpart attack more
+    difficult. We would like to confirm that these changes have been made, by repeating the Ristenpart attack.
+  \item If time permits, we will attempt to defeat Amazon's new countermeasures, or improve on the reliability of the
+    Ristenpart attack, whatever is more appropriate.
 \end{itemize}
@@
 \section{Work Description}
-Here you describe the work you have performed, problems you have solved and methods you have used.
-There is a fine balance between brevity and conciseness and ensuring that other people, if investing the time, would be able to reproduce your results given this description.
+Our planned work consists of two steps: a ``network cartography'' phase which consists of repeating key measurements
+  from the Ristenpart paper, and an ``attack'' phase in which we try to maximize the probability of locating a new
+  Amazon EC2 virtual machine on the same physical host as a target.
+In an effort to accelerate the project, we will contact the authors of the original study and ask to use their source
+  code.
+\subsection{Cartography Phase}
+Our goal for this step is to find the relationship between AWS instance type and IP address.
+We will use the EC2 API to create a large number of virtual machines, record their IP addresses, and then terminate
+  them.
+We will then look for patterns between IP address allocation and instance type.
+We expect to complete this task by March 22.
+\subsection{Attack Phase}
+Following the methods of the Ristenpart paper, we will create a simplistic program to transfer information over a
+cross-VM covert channel between two VMs under our control.
+We will assume that if this method succeeds, the VMs are co-located.
+We will then try to correlate the success of this method with other relationships between the VMs, such as numerically
+  close IP addresses, or identical IP addresses for the VMs' ``domain zero'' host controller.
+We will then use those other relationships to test co-location with arbitrary VMs.
+
+We will first implement a brute force attack that creates a large number of virtual machines with the same parameters,
+  and measure their co-location rate with a number of target VMs.
+We will then implement the Ristenpart et al. improvement that creates the attack virtual machines at approximately the
+  same time as the target VMs, and measure the co-location rate similarly.
+If time permits, we will implement other improvements as well.
+
+We expect to finish this task by March 29.
 

### 2015-04-11 src/seeker.c
make hard drive covert channels work

@@ -1,0 +1,78 @@
+#define _LARGEFILE64_SOURCE
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <unistd.h>
+#include <errno.h>
+#include <time.h>
+#include <signal.h>
+#include <sys/fcntl.h>
+#include <sys/ioctl.h>
+#include <linux/fs.h>
+
+#define BLOCKSIZE 512
+#define TIMEOUT 30
+
+int count;
+time_t start;
+
+void done()
+{
+	time_t end;
+
+	time(&end);
+
+	if (end < start + TIMEOUT) {
+		alarm(1);
+		return;
+	}
+
+	if (count) {
+	  printf("%.2f\n", count / TIMEOUT, 1000.0 * TIMEOUT / count);
+	}
+	exit(EXIT_SUCCESS);
+}
+
+void handle(const char *string, int error)
+{
+	if (error) {
+		perror(string);
+		exit(EXIT_FAILURE);
+	}
+}
+
+int main(int argc, char **argv)
+{
+	char buffer[BLOCKSIZE];
+	int fd, retval;
+	unsigned long numblocks;
+	off64_t offset;
+
+	setvbuf(stdout, NULL, _IONBF, 0);
+
+	if (argc != 2) {
+		printf("Usage: seeker <raw disk device>\n");
+		exit(EXIT_SUCCESS);
+	}
+
+	fd = open(argv[1], O_RDONLY);
+	handle("open", fd < 0);
+
+	retval = ioctl(fd, BLKGETSIZE, &numblocks);
+	handle("ioctl", retval == -1);
+
+	time(&start);
+	srand(start + (time_t)getpid());
+	signal(SIGALRM, &done);
+	alarm(1);
+
+	for (;;) {
+		offset = (off64_t) numblocks * random() / RAND_MAX;
+		retval = lseek64(fd, BLOCKSIZE * offset, SEEK_SET);
+		handle("lseek64", retval == (off64_t) -1);
+		retval = read(fd, buffer, BLOCKSIZE);
+		handle("read", retval < 0);
+		count++;
+	}
+	/* notreached */
+}

### 2015-04-12 data/makehisto.py
add data and stuff

@@ -2,7 +2,7 @@ import statdigest
 import numpy as np
 
 # size of bins in milliseconds
-precision = .2
+precision = .04
 
 def arbRound(num, resolution):
     return round(num / resolution) * resolution

### 2015-04-12 src/makehisto.py

@@ -1,29 +1,0 @@
-import statdigest
-import numpy as np
-
-# size of bins in milliseconds
-precision = .2
-
-def arbRound(num, resolution):
-    return round(num / resolution) * resolution
-
-(zerolist, onelist) = statdigest.getLists()
-
-i = arbRound(min(zerolist), precision)-precision
-max = arbRound(max(onelist), precision)+precision
-bins = []
-while i < max:
-    bins.append(i)
-    i += precision
-bins.append(max)
-(zerohist, _) = np.histogram(zerolist, bins)
-(onehist, _) = np.histogram(onelist, bins)
-
-zerofile = open("zerohist.dat", "w+")
-onefile = open("onehist.dat", "w+")
-
-for (hist, file) in zip([zerohist, onehist],[zerofile, onefile]):
-    file.write("BIN\tCOUNT\n")
-    for i in range(1, len(bins)):
-       file.write("%.1f-%.1f\t%d\n" % (bins[i-1], bins[i], hist[i-1]))
-

### 2015-04-12 src/statdigest.py

@@ -1,15 +1,0 @@
-from statistics import mean, stdev
-import numpy as np
-
-def getLists():
-    return (
-            np.fromfile(open("zerovalues.dat", 'r'), sep='\n'),
-            np.fromfile(open("onevalues.dat", 'r'), sep='\n'))
-
-def compute_stats(list):
-    return (mean(list), min(list), max(list), stdev(list))
-
-print("MEAN\tMIN\tMAX\tSTDEV")
-for l in getLists():
-    print("%.2f\t%.2f\t%.2f\t%.2f" % compute_stats(l))
-

### 2015-04-12 data/makehisto.py
add data from local hard drive test

@@ -1,0 +1,28 @@
+import statdigest
+import numpy as np
+
+# size of bins in milliseconds
+precision = .2
+
+def arbRound(num, resolution):
+    return round(num / resolution) * resolution
+
+(zerolist, onelist) = statdigest.getLists()
+
+i = arbRound(min(zerolist), precision)-precision
+max = arbRound(max(onelist), precision)+precision
+bins = []
+while i < max:
+    bins.append(i)
+    i += precision
+bins.append(max)
+(zerohist, _) = np.histogram(zerolist, bins)
+(onehist, _) = np.histogram(onelist, bins)
+
+outfile = open("hdcc_hist.dat", 'w+')
+
+outfile.write("BIN\t\tZERO\tONE\n")
+for i in range(1, len(bins)):
+   outfile.write("%.1f-%.1f\t%d\t%d\n" %
+            (bins[i-1], bins[i], zerohist[i-1], onehist[i-1]))
+

### 2015-04-13 paper.tex

@@
 \section{Work Description}
-Our planned work consists of two steps: a ``network cartography'' phase which consists of repeating key measurements
-  from the Ristenpart paper, and an ``attack'' phase in which we try to maximize the probability of locating a new
-  Amazon EC2 virtual machine on the same physical host as a target.
-In an effort to accelerate the project, we will contact the authors of the original study and ask to use their source
-  code.
-\subsection{Cartography Phase}
-Our goal for this step is to find the relationship between AWS instance type and IP address.
-We will use the EC2 API to create a large number of virtual machines, record their IP addresses, and then terminate
-  them.
-We will then look for patterns between IP address allocation and instance type.
-We expect to complete this task by March 22.
-\subsection{Attack Phase}
-Following the methods of the Ristenpart paper, we will create a simplistic program to transfer information over a
-cross-VM covert channel between two VMs under our control.
-We will assume that if this method succeeds, the VMs are co-located.
-We will then try to correlate the success of this method with other relationships between the VMs, such as numerically
-  close IP addresses, or identical IP addresses for the VMs' ``domain zero'' host controller.
-We will then use those other relationships to test co-location with arbitrary VMs.
+We reproduced two parts of the Ristenpart attack: first, a ``network cartography'' project that tried to find
+  correlations between AWS instance creation parameters and IP address; and second, a ``covert channel'' approach to
+  verifying co-location.
+Our goal is to determine whether one of our instances shares a virtual machine with an arbitrary AWS instance.
 
-We will first implement a brute force attack that creates a large number of virtual machines with the same parameters,
-  and measure their co-location rate with a number of target VMs.
-We will then implement the Ristenpart et al. improvement that creates the attack virtual machines at approximately the
-  same time as the target VMs, and measure the co-location rate similarly.
-If time permits, we will implement other improvements as well.
+\subsection{Covert Channel}
+Our goal is to be able to discern co-location with an arbitrary instance, but we can begin by attempting to discern
+  co-location with another instance under our control.
+Ristenpart et al. evaluated co-location by attempting to transmit data across a covert channel between two virtual
+  machines --- in their case, by causing seeks and measuring seek times. Our approach is essentially the same.
+A ``sender'' process would continually read values from random locations on the disk to send a one, and do nothing to
+  send a zero.
+A ``receiver'' process would constantly be reading values, and measuring the seek times.
+If the processes shared a hard drive (and thus, a host), the receiver process would be expected to observe consistently
+  higher seek times while the sender is transmitting ones than while the sender is transmitting zeroes.
+Thus, successfully transmitting a message across this secret channel would imply that the virtual machines are
+  co-located.
 
-We expect to finish this task by March 29.
+\subsubsection{Verification on Local Machine}
+Some calibration measurements were taken to make sure that the covert channel approach would work.
+The ``seeker'', a short C program, was adapted from a 2007 Linux Insider article (TODO cite) to provide the seek
+  capability and measurement.
+The seeker read random locations from the disk as frequently as possible for thirty seconds, then reported the mean
+  time required for each reading.
+The seeker was run one hundred times on a personal computer belonging to one of the authors, and the mean seek time
+  from each sample was recorded.
+
+(TODO results from laptop test)
+
+A similar test was performed on the same personal computer that ran two seekers at once.
+One hundred samples were taken from one seeker (the ``reader'').
+The other seeker's samples were discarded.
+
+\subsubsection{Calibration on AWS}
+Once we were satisfied that the seeker was performing as expected, we used it to calibrate the covert channel on a
+  t2.micro AWS instance.
+The one-seeker test and two-seeker test were each run one hundred times, and their sampling distributions were
+  used to set the mean seek time ranges that corresponded to zero and one.
+
+\subsubsection{Discovery of co-located Virtual Machines}
+(TODO \#) (TODO type) instances were set up in the (TODO az) availability zone, spread evenly across two AWS accounts.
+Each virtual machine tried to transmit (TODO \#bits) bits, at a bit rate of one per 30 seconds, to each other virtual
+  machine.
+(TODO \#) virtual machines were able to successfully transmit the message, (TODO \#) of which had one-bit transmission
+  errors in one direction, but not the other.
+We inferred these (TODO \#) pairs to be co-located.
 
 \section{Results}
-\todo{here you will present and discuss your outcomes: implementation results or measurements or other project outcomes}
+\subsection{Covert Channel}
+The tests we ran on our personal computer, described in Section~{mth:localcc} provided information that gave us high
+  confidence in the validity of the results we measured on AWS. The sampling distribution for these tests are shown in
+  Figure~(TODO figure).
+
+Even if the probability distribution of the hard drive seek time was not Gaussian, such a large number of samples were
+  taken that the sampling distribution of means can be treated as Gaussian, according to (TODO who?)'s theorem.
+We have superimposed Gaussian functions on the histogram in Figure~(TODO figure) to demonstrate our analysis.
+
+The distributions are very narrow and widely separated, giving us high confidence in the accuracy of our message
+  transmission.
+The sampling distribution of mean seek time in the single-seeker test (transmitting zeroes) was centered on (TODO \#)
+  (TODO milliseconds), with a standard deviation of (TODO \#)(TODO milliseconds).
+For the double-seeker test (transmitting ones), the center was (TODO milliseconds) with a standard deviation of (TODO
+  milliseconds).
+Based on these calibration figures, we could read a mean seek time between (TODO milliseconds) and (TODO milliseconds)
+  as a zero, and a mean seek time between (TODO milliseconds) and (TODO milliseconds), and accomplish an error rate of
+  (TODO error rate) percent.
+
+Our first calibration test on AWS produced an unexpected result; the first \num{53} samples of the one-seeker test gave
+  a mean seek time between \SI{0.29}{ms} and \SI{0.31}{ms}, while the last \num{46} were all over \SI{36}{ms}.
+Sample \num{54} gave a mean of \SI{0.81}{ms}, indicating that toward the end of that sample, some condition changed
+  that affected the seek time of the VM.
+The two-seeker test, which ran after the one-seeker test, showed sample means in the \SI{70}{ms} range.
+
+Reducing the number of samples in the test to (TODO number) eliminated the shift as a factor.
+However, the test results reported extremely fast seeks, with no discernible difference between the one-seeker test and
+  the two-seeker tests.
+Details are shown in Figure~(TODO ref).
+
+(TODO figure)
+
+We believe that EC2 instances are now most often backed by Solid State Disks, which do not exhibit the seek-time
+  properties we have been trying to exploit.
+Instead of requiring an arm to physically rotate into place over a magnetic platter, SSDs provide fast, constant-time
+  random access, a property that eliminates the covert channel we were trying to communicate on.
+We believe the sudden change in seek times during the first experiment was caused by our virtual machine moving suddenly
+
+The calibration tests we ran on AWS provided a window of (TODO milliseconds)-(TODO milliseconds) for zeroes, and (TODO
+  milliseconds)-(TODO milliseconds) for ones, with a similar error rate.
+Our VM matching experiment transmitted (TODO \#bits) bits between co-located virtual machines and observed
+  (TODO \#errors) errors, a rate of (TODO error rate) percent; the deviation from the expected error rate is not
+  significant ($p = (TODO p-value)$).
 

### 2015-04-13 paper.tex
paper updates

@@
 \usepackage{booktabs,threeparttable,multirow}
+\usepackage{siunitx}
+\usepackage{pgfplots}
 
@@
 We would like to discern whether Amazon AWS has implemented any countermeasures against co-location attacks since the
-first successful cloud co-location six years ago; we will repeat their procedure in an attempt to reproduce their
-results. If time permits, we will try to improve on their attack.
+  first successful cloud co-location six years ago.
+We repeated one aspect of a previous successful attack, and found that, surprisingly, the same attack vector continues
+  to be plausible.
 \end{abstract}
@@
 \section{Motivation}
-\todo{Know what you want to do and why that is interesting (maybe with bullet points). But do not write this section until you know what you actually have done so that the motivation fits your work.}
 
@@
   machines --- in their case, by causing seeks and measuring seek times. Our approach is essentially the same.
-A ``sender'' process would continually read values from random locations on the disk to send a one, and do nothing to
+A ``sender'' process continually reads values from random locations on the disk to send a one, and does nothing to
   send a zero.
@@
 Some calibration measurements were taken to make sure that the covert channel approach would work.
-The ``seeker'', a short C program, was adapted from a 2007 Linux Insider article (TODO cite) to provide the seek
+The ``seeker'', a short C program, was adapted from a 2007 Linux Insight article \cite{seeker07} to provide the seek
   capability and measurement.
@@
 
-(TODO results from laptop test)
-
 A similar test was performed on the same personal computer that ran two seekers at once.
-One hundred samples were taken from one seeker (the ``reader'').
+One hundred samples were taken from one seeker (the ``receiver'').
 The other seeker's samples were discarded.
@@
 
-\subsubsection{Discovery of co-located Virtual Machines}
-(TODO \#) (TODO type) instances were set up in the (TODO az) availability zone, spread evenly across two AWS accounts.
-Each virtual machine tried to transmit (TODO \#bits) bits, at a bit rate of one per 30 seconds, to each other virtual
-  machine.
-(TODO \#) virtual machines were able to successfully transmit the message, (TODO \#) of which had one-bit transmission
-  errors in one direction, but not the other.
-We inferred these (TODO \#) pairs to be co-located.
-
 \section{Results}
@@
 The tests we ran on our personal computer, described in Section~{mth:localcc} provided information that gave us high
-  confidence in the validity of the results we measured on AWS. The sampling distribution for these tests are shown in
-  Figure~(TODO figure).
+  confidence in the validity of the results we measured on AWS.
 
 Even if the probability distribution of the hard drive seek time was not Gaussian, such a large number of samples were
-  taken that the sampling distribution of means can be treated as Gaussian, according to (TODO who?)'s theorem.
-We have superimposed Gaussian functions on the histogram in Figure~(TODO figure) to demonstrate our analysis.
+  taken that the sampling distribution of means can be treated as Gaussian, according to the Central Limit Theorem.
 
@@
   transmission.
-The sampling distribution of mean seek time in the single-seeker test (transmitting zeroes) was centered on (TODO \#)
-  (TODO milliseconds), with a standard deviation of (TODO \#)(TODO milliseconds).
-For the double-seeker test (transmitting ones), the center was (TODO milliseconds) with a standard deviation of (TODO
-  milliseconds).
-Based on these calibration figures, we could read a mean seek time between (TODO milliseconds) and (TODO milliseconds)
-  as a zero, and a mean seek time between (TODO milliseconds) and (TODO milliseconds), and accomplish an error rate of
-  (TODO error rate) percent.
+The sampling distribution of mean seek time in the single-seeker test (transmitting zeroes) was centered on
+  \SI{13.36}{ms}, with a standard deviation of \SI{0.36}{ms}.
+For the double-seeker test (transmitting ones), the center was \SI{24.08}{ms} with a standard deviation of
+  \SI{1.42}{ms}.
+Based on these calibration figures, we could read a mean seek time between \SI{11.56}{ms} and \SI{15.16}{ms}
+  as a zero, and a mean seek time between \SI{16.98}{ms} and \SI{31.18}{ms} as a one, and accomplish an error rate of
+  \SI{3e-7} errors per bit transmitted.
 
@@
 
-Reducing the number of samples in the test to (TODO number) eliminated the shift as a factor.
-However, the test results reported extremely fast seeks, with no discernible difference between the one-seeker test and
-  the two-seeker tests.
-Details are shown in Figure~(TODO ref).
-
-(TODO figure)
-
-We believe that EC2 instances are now most often backed by Solid State Disks, which do not exhibit the seek-time
-  properties we have been trying to exploit.
-Instead of requiring an arm to physically rotate into place over a magnetic platter, SSDs provide fast, constant-time
-  random access, a property that eliminates the covert channel we were trying to communicate on.
-We believe the sudden change in seek times during the first experiment was caused by our virtual machine moving suddenly
-
-The calibration tests we ran on AWS provided a window of (TODO milliseconds)-(TODO milliseconds) for zeroes, and (TODO
-  milliseconds)-(TODO milliseconds) for ones, with a similar error rate.
-Our VM matching experiment transmitted (TODO \#bits) bits between co-located virtual machines and observed
-  (TODO \#errors) errors, a rate of (TODO error rate) percent; the deviation from the expected error rate is not
-  significant ($p = (TODO p-value)$).
+Reducing the number of samples in the test to ten eliminated the shift as a factor.
+The EC2 tests exhibited a much faster overall seek time than the local tests.
+However, the one-seek and two-seek tests still exhibited dramatically different results, with the all one-seek means
+  being between \SI{0.31}{ms} and \SI{0.33}{ms}, and all two-seek times between \SI{0.58}{ms} and \SI{0.61}{ms}.
 
 \section{Conclusion}
-\todo{TBD last}
+Storage drive based covert channels still represent a viable way to verify instance co-location on Amazon EC2.
+Future work would consist of running a large-scale experiment on EC2, where instances would try to communicate with
+  other instances on the same host by transmitting and receiving on the channel that we have demonstrated again here.
+Correlating these co-located instances with publicly available network properties would enable us to predict
+  co-location of arbitrary VMs.
 
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
 

### 2015-04-13 data/makehisto.py

@@ -23,6 +23,6 @@ outfile = open("hdcc_hist.dat", 'w+')
 
 outfile.write("BIN\t\tZERO\tONE\n")
 for i in range(1, len(bins)):
-   outfile.write("%.1f-%.1f\t%d\t%d\n" %
+   outfile.write("%.2f-%.2f\t%d\t%d\n" %
             (bins[i-1], bins[i], zerohist[i-1], onehist[i-1]))
 

### 2015-04-13 paper.tex
finish some stuff-

@@
 \IEEEoverridecommandlockouts
+
+\input{preamble}
+
 \begin{document}
@@
 \maketitle
+\nocite{*}
 
@@
   \item We expect that Amazon has since adjusted their VM location algorithms to make the Ristenpart attack more
-    difficult. We would like to confirm that these changes have been made, by repeating the Ristenpart attack.
-  \item If time permits, we will attempt to defeat Amazon's new countermeasures, or improve on the reliability of the
-    Ristenpart attack, whatever is more appropriate.
+    difficult. We would like to assess the current viability of the Ristenpart attack on today's AWS.
 \end{itemize}
@@
 \section{Background}\label{sec:background}
-\todo{You should find and describe related work early on. Know what other people have done.}
+The idea of performing cryptanalytic attacks against co-located virtual machines was discussed by Ristenpart, Shacham,
+  and Savage in 2009 \cite{ristenpart09}.
+They measure co-location success rates in Amazon Web Services (AWS), and suggest some potential attacks against
+  co-located VMs.
+Zhang \cite{zhang2014cross} et al. presented a cross-VM attack using Ristenpart's attack which was able to recover an El Gamal key across
+Xen VMs.
+Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoquis2014wait} then presented a practical, fast, and realistic full key
+  recovery against AES across virtual machines, based on Bernstein's \cite{bernstein2005cache} attack.
+
+Research into co-location attacks and cross-VM attacks continues to be an active research area.
 
@@
 \bibliographystyle{IEEEtran}
-
+\bibliography{bibliography}
 \end{document}

### 2015-04-13 paper.tex
merge the paper

@@
 \usepackage{pgfplots}
+\usepackage{graphicx}
+\graphicspath{ {images/} }
 
@@
 The idea of performing cryptanalytic attacks against co-located virtual machines was discussed by Ristenpart, Shacham,
-  and Savage in 2009 \cite{ristenpart09}.
+  and Savage in 2009 \cite{ristenpart2009cloud}.
 They measure co-location success rates in Amazon Web Services (AWS), and suggest some potential attacks against
@@
 Xen VMs.
-Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoquis2014wait} then presented a practical, fast, and realistic full key
+Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoqui2014wait} then presented a practical, fast, and realistic full key
   recovery against AES across virtual machines, based on Bernstein's \cite{bernstein2005cache} attack.
@@
 
-\subsection{Covert Channel}
+\subsection{Network Cartography}
+In this section, we create a “map” of instance placements by recording the private IP address of the instances after placing them.
+Our hope is this will give insight into how target instances are placed.
+This will help in the attempt to establish co-residence with these target instances.
+
+From published information about EC2, we know that each availability zone has its own dedicated hardware. 
+Using this and the cartography information in Ristenpart, we hypothesize that each zone will have its own 
+range of internal IP addresses. 
+We also hypothesize that instances of the same type would fall in the same internal IP address ranges. 
+If these hypotheses are correct, an attacker could achieve co-location much more easily by finding the internal 
+IP address of the target and launching instances with similar parameters. 
+Amazon’s DNS service can be used to obtain this address from the public IP address of the target.
+
+We test our theory by launching a number of instances with different types in each availability zone. 
+We then attempt to create a set of rules that describe the likely availability zone and instance type of a given 
+private IP address.  
+With the data collected, we create a “map” of IP addresses and instance parameters that can be used to determine 
+a target’s parameters.
+
+All instances were launched in US West 2 zones a, b, and c. The t2.micro, small, and medium instance types 
+were used. 
+Of the 1011 instances launched, all fell into the same /16 prefix (172.31.0.0). 
+Instances were launched 
+in groups of 3 or 5 of each type. 
+We observed that instances of the same type launched sequentially had consecutive IP addresses, which is called 
+sequential placement locality.
+
+While it was fairly likely that instances launched in the same zone would be placed together based on previous 
+work and knowledge of the inner workings of EC2, we were surprised to see that different instances types were launched 
+very close to each other (within the same /24 prefix). 
+Of the 48 /24 prefixes observed across the three availability zones, the vast majority contained more than one 
+instance type. 
+Additionally, a small number of m3.medium, large, and xlarge instances were launched which all fell into the same 
+/20 prefix as other instances in that zone.
+
+The following plots show the results of the data collected. 
+IP addresses are represented by their integer representation minus 2887712768, which represents the first address 
+(172.31.0.0). 
+Zone 2c occupied the 172.31.0.0/20 prefix, Zone 2a occupied 172.31.16.0/20, and Zone 2b occupied 172.31.32.0/20. 
+In (this figure), the zones are numbered one to three, with a being one.
+%\begin{figure}[h]
+%\includegraphics[scale=.5]{zoneIP}
+%\caption{Plot of private IP address vs availability zone}
+%\end{figure}
+In (these figures), the distribution of instance types in each zone is showed. The instances are numbered from 
+one to three, for t2.micro, t2.small, and t2.medium instances respectively.
+%\begin{figure}[h]
+%\includegraphics[scale=.5]{2cTypeIP}
+%\caption{Plot of private IP address vs instance type in Zone 2c}
+%\end{figure}
+%\begin{figure}[h]
+%\includegraphics[scale=.5]{2aTypeIP}
+%\caption{Plot of private IP address vs instance type in Zone 2a}
+%\end{figure}
+%\begin{figure}[h]
+%\includegraphics[scale=.5]{2bTypeIP}
+%\caption{Plot of private IP address vs instance type in Zone 2b}
+%\end{figure}
 Our goal is to be able to discern co-location with an arbitrary instance, but we can begin by attempting to discern
@@
   send a zero.
-A ``receiver'' process would constantly be reading values, and measuring the seek times.
+A ``receiver'' process is constantly reading values, and measuring the seek times.
 If the processes shared a hard drive (and thus, a host), the receiver process would be expected to observe consistently
@@
 
+Our first calibration test on AWS produced an unexpected result; the first \num{53} samples of the one-seeker test gave
+  a mean seek time between \SI{0.29}{ms} and \SI{0.31}{ms}, while the last \num{46} were all over \SI{36}{ms}.
+Sample \num{54} gave a mean of \SI{0.81}{ms}, indicating that toward the end of that sample, some condition changed
+  that affected the seek time of the VM.
+The two-seeker test, which ran after the one-seeker test, showed sample means in the \SI{70}{ms} range.
+
+Reducing the number of samples in the test to ten eliminated the shift as a factor.
+The EC2 tests exhibited a much faster overall seek time than the local tests.
+However, the one-seek and two-seek tests still exhibited dramatically different results, with the all one-seek means
+  being between \SI{0.31}{ms} and \SI{0.33}{ms}, and all two-seek times between \SI{0.58}{ms} and \SI{0.61}{ms}.
+
 \section{Conclusion}

### 2015-04-14 paper.tex
writing about cartography

@@
 \usepackage{booktabs,threeparttable,multirow}
-\usepackage{siunitx}
-\usepackage{pgfplots}
 \usepackage{graphicx}
@@
 \IEEEoverridecommandlockouts
-
-\input{preamble}
-
 \begin{document}
@@
 \maketitle
-\nocite{*}
 
@@
 We would like to discern whether Amazon AWS has implemented any countermeasures against co-location attacks since the
-  first successful cloud co-location six years ago.
-We repeated one aspect of a previous successful attack, and found that, surprisingly, the same attack vector continues
-  to be plausible.
+first successful cloud co-location six years ago; we will repeat their procedure in an attempt to reproduce their
+results. If time permits, we will try to improve on their attack.
 \end{abstract}
@@
 \section{Motivation}
+\todo{Know what you want to do and why that is interesting (maybe with bullet points). But do not write this section until you know what you actually have done so that the motivation fits your work.}
 
@@
   \item We expect that Amazon has since adjusted their VM location algorithms to make the Ristenpart attack more
-    difficult. We would like to assess the current viability of the Ristenpart attack on today's AWS.
+    difficult. We would like to confirm that these changes have been made, by repeating the Ristenpart attack.
+  \item If time permits, we will attempt to defeat Amazon's new countermeasures, or improve on the reliability of the
+    Ristenpart attack, whatever is more appropriate.
 \end{itemize}
@@
 \section{Background}\label{sec:background}
-The idea of performing cryptanalytic attacks against co-located virtual machines was discussed by Ristenpart, Shacham,
-  and Savage in 2009 \cite{ristenpart2009cloud}.
-They measure co-location success rates in Amazon Web Services (AWS), and suggest some potential attacks against
-  co-located VMs.
-Zhang \cite{zhang2014cross} et al. presented a cross-VM attack using Ristenpart's attack which was able to recover an El Gamal key across
-Xen VMs.
-Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoqui2014wait} then presented a practical, fast, and realistic full key
-  recovery against AES across virtual machines, based on Bernstein's \cite{bernstein2005cache} attack.
-
-Research into co-location attacks and cross-VM attacks continues to be an active research area.
+\todo{You should find and describe related work early on. Know what other people have done.}
 
@@
 %\end{figure}
+\subsection{Covert Channel}
 Our goal is to be able to discern co-location with an arbitrary instance, but we can begin by attempting to discern
@@
   machines --- in their case, by causing seeks and measuring seek times. Our approach is essentially the same.
-A ``sender'' process continually reads values from random locations on the disk to send a one, and does nothing to
+A ``sender'' process would continually read values from random locations on the disk to send a one, and do nothing to
   send a zero.
-A ``receiver'' process is constantly reading values, and measuring the seek times.
+A ``receiver'' process would constantly be reading values, and measuring the seek times.
 If the processes shared a hard drive (and thus, a host), the receiver process would be expected to observe consistently
@@
 Some calibration measurements were taken to make sure that the covert channel approach would work.
-The ``seeker'', a short C program, was adapted from a 2007 Linux Insight article \cite{seeker07} to provide the seek
+The ``seeker'', a short C program, was adapted from a 2007 Linux Insider article (TODO cite) to provide the seek
   capability and measurement.
@@
 
+(TODO results from laptop test)
+
 A similar test was performed on the same personal computer that ran two seekers at once.
-One hundred samples were taken from one seeker (the ``receiver'').
+One hundred samples were taken from one seeker (the ``reader'').
 The other seeker's samples were discarded.
@@
 
+\subsubsection{Discovery of co-located Virtual Machines}
+(TODO \#) (TODO type) instances were set up in the (TODO az) availability zone, spread evenly across two AWS accounts.
+Each virtual machine tried to transmit (TODO \#bits) bits, at a bit rate of one per 30 seconds, to each other virtual
+  machine.
+(TODO \#) virtual machines were able to successfully transmit the message, (TODO \#) of which had one-bit transmission
+  errors in one direction, but not the other.
+We inferred these (TODO \#) pairs to be co-located.
+
 \section{Results}
@@
 The tests we ran on our personal computer, described in Section~{mth:localcc} provided information that gave us high
-  confidence in the validity of the results we measured on AWS.
+  confidence in the validity of the results we measured on AWS. The sampling distribution for these tests are shown in
+  Figure~(TODO figure).
 
 Even if the probability distribution of the hard drive seek time was not Gaussian, such a large number of samples were
-  taken that the sampling distribution of means can be treated as Gaussian, according to the Central Limit Theorem.
+  taken that the sampling distribution of means can be treated as Gaussian, according to (TODO who?)'s theorem.
+We have superimposed Gaussian functions on the histogram in Figure~(TODO figure) to demonstrate our analysis.
 
@@
   transmission.
-The sampling distribution of mean seek time in the single-seeker test (transmitting zeroes) was centered on
-  \SI{13.36}{ms}, with a standard deviation of \SI{0.36}{ms}.
-For the double-seeker test (transmitting ones), the center was \SI{24.08}{ms} with a standard deviation of
-  \SI{1.42}{ms}.
-Based on these calibration figures, we could read a mean seek time between \SI{11.56}{ms} and \SI{15.16}{ms}
-  as a zero, and a mean seek time between \SI{16.98}{ms} and \SI{31.18}{ms} as a one, and accomplish an error rate of
-  \SI{3e-7} errors per bit transmitted.
+The sampling distribution of mean seek time in the single-seeker test (transmitting zeroes) was centered on (TODO \#)
+  (TODO milliseconds), with a standard deviation of (TODO \#)(TODO milliseconds).
+For the double-seeker test (transmitting ones), the center was (TODO milliseconds) with a standard deviation of (TODO
+  milliseconds).
+Based on these calibration figures, we could read a mean seek time between (TODO milliseconds) and (TODO milliseconds)
+  as a zero, and a mean seek time between (TODO milliseconds) and (TODO milliseconds), and accomplish an error rate of
+  (TODO error rate) percent.
 
@@
 
-Reducing the number of samples in the test to ten eliminated the shift as a factor.
-The EC2 tests exhibited a much faster overall seek time than the local tests.
-However, the one-seek and two-seek tests still exhibited dramatically different results, with the all one-seek means
-  being between \SI{0.31}{ms} and \SI{0.33}{ms}, and all two-seek times between \SI{0.58}{ms} and \SI{0.61}{ms}.
+Reducing the number of samples in the test to (TODO number) eliminated the shift as a factor.
+However, the test results reported extremely fast seeks, with no discernible difference between the one-seeker test and
+  the two-seeker tests.
+Details are shown in Figure~(TODO ref).
 
-Our first calibration test on AWS produced an unexpected result; the first \num{53} samples of the one-seeker test gave
-  a mean seek time between \SI{0.29}{ms} and \SI{0.31}{ms}, while the last \num{46} were all over \SI{36}{ms}.
-Sample \num{54} gave a mean of \SI{0.81}{ms}, indicating that toward the end of that sample, some condition changed
-  that affected the seek time of the VM.
-The two-seeker test, which ran after the one-seeker test, showed sample means in the \SI{70}{ms} range.
+(TODO figure)
 
-Reducing the number of samples in the test to ten eliminated the shift as a factor.
-The EC2 tests exhibited a much faster overall seek time than the local tests.
-However, the one-seek and two-seek tests still exhibited dramatically different results, with the all one-seek means
-  being between \SI{0.31}{ms} and \SI{0.33}{ms}, and all two-seek times between \SI{0.58}{ms} and \SI{0.61}{ms}.
+We believe that EC2 instances are now most often backed by Solid State Disks, which do not exhibit the seek-time
+  properties we have been trying to exploit.
+Instead of requiring an arm to physically rotate into place over a magnetic platter, SSDs provide fast, constant-time
+  random access, a property that eliminates the covert channel we were trying to communicate on.
+We believe the sudden change in seek times during the first experiment was caused by our virtual machine moving suddenly
+
+The calibration tests we ran on AWS provided a window of (TODO milliseconds)-(TODO milliseconds) for zeroes, and (TODO
+  milliseconds)-(TODO milliseconds) for ones, with a similar error rate.
+Our VM matching experiment transmitted (TODO \#bits) bits between co-located virtual machines and observed
+  (TODO \#errors) errors, a rate of (TODO error rate) percent; the deviation from the expected error rate is not
+  significant ($p = (TODO p-value)$).
 
 \section{Conclusion}
-Storage drive based covert channels still represent a viable way to verify instance co-location on Amazon EC2.
-Future work would consist of running a large-scale experiment on EC2, where instances would try to communicate with
-  other instances on the same host by transmitting and receiving on the channel that we have demonstrated again here.
-Correlating these co-located instances with publicly available network properties would enable us to predict
-  co-location of arbitrary VMs.
+\todo{TBD last}
 
-\bibliographystyle{IEEEtran}
-\bibliography{bibliography}
+%\bibliographystyle{IEEEtran}
+
 \end{document}
